---
title: "Exercise_B"
format: html
editor: visual
---

We will need the following to functions fromÂ [Exercise A](https://computationalmovementanalysis.github.io/FS24/Week2/3_exercise_a.html):

```{r}
library("readr")
library("sf")
library("dplyr")

difftime_secs <- function(x, y){
  as.numeric(difftime(x, y, units = "secs"))
}

distance_by_element <- function(later, now){
  as.numeric(
    st_distance(later, now, by_element = TRUE)
  )
}
```

We can then import the data. We can discard all columns with the exception ofÂ `DatetimeUTC`.

```{r}
caro <- read_delim("caro60.csv", ",") |>
  st_as_sf(coords = c("E","N"), crs = 2056) |> 
  select(DatetimeUTC)
```

# Task 1: Calculate speed at scale 1:

In our first scale, we will assume a sampling windowÂ ğ‘¤Â of 120 seconds. This conveniently means that for every location, you can use the previous and next location to calculate speed. Try to implement this in R.

First Calculate timelag and steplength like in the previous Exercise

```{r}
caro <- mutate(caro, timelag = as.numeric(difftime_secs(lead(DatetimeUTC, n=1), lag(DatetimeUTC, n=1))))

caro <- mutate(caro, steplength = as.numeric(distance_by_element(lead(geometry, n=1), lag(geometry, n=1))))

```

Then use these variables to calculate the speed like in Exercise A

```{r}
caro <- mutate(caro, speed = steplength/timelag)

head(caro)
```

# Task 2: Calculate speed at scale 2

To compare the effect of different sampling intervals, Laube and Purves ([2011](https://computationalmovementanalysis.github.io/FS24/90_references.html#ref-laube2011)) calculated speed at different scales (i.e.Â different sampling windows ğ‘¤).

In the previous task, we assumed a ğ‘¤=120ğ‘ . In this task, try to implement ğ‘¤=240ğ‘  (see [FigureÂ 8.2](https://computationalmovementanalysis.github.io/FS24/Week2/3_exercise_b.html#fig-laube2011b)).

-   Tip: Use the `n =` parameter in `lead`/`lag` to increase the offset.

-   Store values *timelag*, *steplength* and *speed* in the columns `timelag2`, `steplength2` and `speed2` to distinguish them from the values from scale 1

First Calculate timelag and steplength like in the previous Exercise just with another scale:

```{r}
caro <- mutate(caro, timelag2 = as.numeric(difftime_secs(lead(DatetimeUTC, n=2), lag(DatetimeUTC, n=2))))

caro <- mutate(caro, steplength2 = as.numeric(distance_by_element(lead(geometry, n=2), lag(geometry, n=2))))
```

Then use these variables to calculate the speed like in Exercise A

```{r}
caro <- mutate(caro, speed2 = steplength2/timelag2)

head(caro)
```

# Task 3: Calculate speed at scale 3:

Redo the previous task withÂ w = 480s

```{r}
caro <- mutate(caro, timelag3 = as.numeric(difftime_secs(lead(DatetimeUTC, n=4), lag(DatetimeUTC, n=4))))

caro <- mutate(caro, steplength3 = as.numeric(distance_by_element(lead(geometry, n=4), lag(geometry, n=4))))
```

then speed:

```{r}
caro <- mutate(caro, speed3 = steplength3/timelag3)

head(caro)
```

# **Task 4: Compare speed across scales**

We now have a dataframe with three different speed values per sample, corresponding to the different scales / sampling windows (ğ‘¤1=120ğ‘ , ğ‘¤2=240ğ‘  and ğ‘¤3=480ğ‘ ). It would now be interesting to compare these measurements and see our results correspond to those of Laube and Purves ([2011](https://computationalmovementanalysis.github.io/FS24/90_references.html#ref-laube2011)). In their experiments, the authors observe:

> -   A steady decrease in median speed as the temporal analysis scale increases;
>
> -   A decrease in the overall variance in speed as the temporal scale increases;
>
> -   Lower minimum values at the shortest temporal scales;

The authors visualize these observations using box plots. To to the same, we need to process our data slightly. Currently, our data looks like this:

```{r}
caro |> 
  st_drop_geometry() |> 
  select(DatetimeUTC, speed, speed2, speed3)
```

We can make a box plot of a single column using ggplot2:

```{r}
library(ggplot2)

ggplot(caro, aes(y = speed)) + 
    # we remove outliers to increase legibility, analogue
  # Laube and Purves (2011)
  geom_boxplot(outlier.shape = NA)
```

However, if we want to compareÂ `speed`Â withÂ `speed2`Â andÂ `speed3`, we need need aÂ *long*Â table rather thanÂ *wide*Â one (which is what we currently have). To make our tableÂ *long*, we can use the functionÂ `pivot_longer`Â fromÂ `tidyr`:

```{r}
library(tidyr)

# before pivoting, let's simplify our data.frame
caro2 <- caro |> 
  st_drop_geometry() |> 
  select(DatetimeUTC, speed, speed2, speed3)

caro_long <- caro2 |> 
  pivot_longer(c(speed, speed2, speed3))
  
head(caro_long)
```

```{r}
ggplot(caro_long, aes(name, value)) +
  # we remove outliers to increase legibility, analogue
  # Laube and Purves (2011)
  geom_boxplot(outliers = FALSE)+
  theme_minimal()
 
```
